An Alternative Geometric Interpretation of the Riemann-Stieltjes Integral

### 2017-11-14 paper.tex
added new changes to paper

\title{An Alternative Geometric Interpretation of the Riemann-Stieltjes Integral}
\markright{Abbreviated Article Title}
\author{Woodrow Wilson and Herbert Hoover}

@@

\begin{abstract}
It is desirable to have geometric interpretations for difficult mathematical ideas. In this paper we present an intuitive geometric interpretation of the Riemann-Stieltjes integral. It can be 
summarized as follows: the Riemann-Stieltjes integral can be inerpreted as an infinite sum of infinitesimally small non-rectangular 
integration strips. This interpretation is surprisingly similar to the geometric interpretation of the ordinary Riemann integral. 
It is therefore likely that this geometric interpretation would reduce the large jump 
students experience when they first encounter the Riemann-Stieltjes integral (i.e. make the 
Riemann-Stieltjes integral less abstract).
\end{abstract}

@@

\noindent
Ackermann et al. utilized the idea of using non-rectangular integration strips to perform integration
and in doing so developed the Cavalieri integral \cite{ackermann2012}. The area of a non-rectangular integration strip is computed via 
Cavalieri's principle \cite{andersen1985,eves1991, malik1984, wildberger2002, young1985}. They also showed that a Cavalieri integral can easily be converted to a Riemann-Stieltjes integral. The 
reverse operation was however not described in \cite{ackermann2012}. In this paper we address this shortcoming. This 
reverse operation actually allows us to contrive a novel geometric interpretation of the Riemann-Stieltjes integral.
This geometric interpretation is: the Riemann-Stieltjes integral can be interpreted as an infinite sum of infinitesimally small
non-rectangular integration strips. Interestingly enough, this interpretation of the Riemann-Stieltjes integral is very similar 
to the geometric interpretation of the Riemann integral. The difference being: in the case of the Riemann integral we employ 
rectangular instead of non-rectangular integration strips. This geometric interpretation has been quite useful while 
teaching the ordinary Riemann integral to undergraduate students \cite{jones2015,sealey2014}. It, therefore, seems likely that undergraduate students would find the non-rectangular geometric interpretation
of the Riemann-Stieltjes integral just as useful. It is also interesting to mention, that there is a few papers in literature that proposes to first 
teach simpler integral definitions (which are easier to grasp) before introducing students to the Riemann integral \cite{czarnocha2001,prabhu2008,ruffa2002,zaskis2014}. It therefore seems plausible that
students will find it useful to first understand the Cavalieri integral before the Riemann-Stieltjes integral is taught to them.

@@

\begin{definition}[Definition of integral]
Define a partition $\mathcal{P}$ of $[a',b']$ to be a set of points $x_0, x_1,\cdots,x_n$, 
where $a' = x_0 \leq x_1 \leq \cdots \leq x_n = b'$. For each partition 
$\mathcal{P}$ of $[a',b']$ write $\Delta g_k = g(x_k)-g(x_{k-1})$. Let 
$M_k = \max\{f(x),x_{k-1}\leq x \leq x_{k}\}$, $m_k = \max\{f(x),x_{k-1}\leq x \leq x_{k}\}$, and set 
\begin{equation}
\label{ref:S_up}
\mathcal{S}_U(\mathcal{P},f,g) = \sum_{k=1}^n M_k \Delta g_k, 
\end{equation}
and
\begin{equation}
\label{ref:S_low}
\mathcal{S}_L(\mathcal{P},f,g) = \sum_{k=1}^n m_k \Delta g_k. 
\end{equation}
The sums in Eq.~\eqref{ref:S_up} and Eq.~\eqref{ref:S_low} are respectively called the the upper and lower Riemann-Stieltjes sums.
If there is a unique number $I$ that satisfies the inequality $\mathcal{S}_L(\mathcal{P},f,g)\leq I \leq \mathcal{S}_U(\mathcal{P},f,g)$ for all 
partitions $P$ of $[a',b']$, then $f$ is Riemann integrable with respect to $g$ on $[a',b']$. Moreover, $I$ is called the Riemann-Stieltjes integral of $f$ from $a'$ to $b'$ and is denoted by
\begin{equation}
\int_{a'}^{b'} f(x) dg(x). 
\end{equation}
\end{definition}

@@

\begin{theorem}
If $a'<s<b'$, $f$ is continuous at $s$, and 
\begin{equation}
g(x) = 
\begin{cases}
0 & \textrm{if}~x\leq s\\
1 & \textrm{if}~x > s
\end{cases}
\end{equation}
then $\int_a^b f dg(x) = f(s)$.
\end{theorem}

@@

\section{3D Geometric Interpretation}
Bullock proposed one possible geometric interpretation of the Riemann-Stieltjes integral. 
The basic premise behind his approach is to represent $f(x)$ and $g(x)$ as three-dimensional sheets. 
The three dimensional coordinate system which Bullock proposes to realize this representation is depicted in 
Fig.~\ref{fig:bullock}. The $f$-sheet is constructed by simply extending the curve $f$ which resides in the $fx$-plane 
along the $g$-direction. The $g$-sheet is constructed similarly. Note that, since $f$ is independent of $g$, the $f$-sheet remains 
constant along the $g$-direction. Similarly, the $g$-sheet remains constant along the $f$-direction. 
A new sheet, known simply as the ``fence'', can be created by chopping off any part of the $g$-sheet that protrudes above the 
$f$-sheet and below the $gx$-plane. This sheet can now be projected onto the $fg$-plane. The area of the region that is formed 
via this projection can be interpreted as the value which is obtained when the Riemann-Stieltjes integral is evaluated.

\noindent
Let us now try to make the above ideas more concrete by applying them to an example. Let $f(x)=x^2$ and 
$g(x)=x$. The sheets associated with these two functions are depicted in Fig.~\ref{fig:bullock}. 
The $f$-sheet is depicted in blue, while the $g$-sheet is red. The ``fence'' is depicted in green. The region
obtained by projecting the ``fence'' on to the $fg$-plane is depicted in black.

\noindent
It is interesting to note that Bullock follows a similar route than we do in terms of developing his geometrical framework.
He starts of with quite strict requirements for $f$ ang $g$ and relaxes more and more of these requirements in latter 
examples.

\begin{figure}[htb]
\centering
\includegraphics[width=1.0\textwidth]{3D_final-crop.pdf}
\caption{An example of using Bullocks' geometric framework to interpret the Riemann-Stieltjes integral. In this example 
$f(x)$ is represented by the blue sheet, $g(x)$ is represented by the red sheet. Moreover, $f(x)=x^2$ and $g(x)=x$. The green sheet, simply known as the ``fence'', is formed 
by chopping of any part of the $g$-sheet which protrudes above the $f$-sheet and below the $gx$-plane. Projecting the ``fence''
onto the $fg$-plane produces the visible gray region in the $fg$-plane. The area of this gray region in the $fg$-plane is equal to the Riemann-Stieltjes integral:
$\int_{0}^{3} f(x) dg(x)$.}
\label{fig:bullock}
\end{figure}

@@

\begin{definition}
A continuous real-valued function $a(y)$ is called a translational function with respect to a continuous real-valued function $f(x)$ on the interval $[a,b]$ if 
$\{x\in\mathbb{R}|a\circ f(x) + z = x\}$ is a singleton, for every $z\in[0,b-a]$ and $a(0) = a$.
\end{definition}

\begin{definition}
The mapping $h : [a, b] \rightarrow [a',b']$, which maps $x_i^1 \in [a, b]$ to $x_i^2 \in [a',b']$, is defined as
$h(x_i^1) :=$ $\{x_i^2 \in [a' ,b'] | a\circ f(x_i^2) + [x_i^1 - a] = x_i^2$ , $a = a(0)\}$, which we will refer to
as the transformation function. 
\end{definition}

@@

\begin{definition}
The mapping $g:[a′, b′] \rightarrow [a, b]$, which maps $x_i^2 \in [a' , b']$ to $x_i^1\in [a, b]$,
is defined as $g(x_i^2) := x_i^2 - a \circ f (x_i^2) + a$, which we will refer to as the inverse
transformation function. 
\end{definition}

@@

\begin{definition}
Define a partition $\mathcal{P}_1$ of $[a,b]$ to be a set of points $x_0, x_1,\cdots,x_n$, 
where $a = x_0^1 \leq x_1^1 \leq \cdots \leq x_n^1 = b$. For each partition 
$\mathcal{P}_1$ of $[a,b]$ write $\Delta x_k = x_k-x_{k-1}$. 
%Since both boundaries of any integration strip are necessarily
%translations of the translational function $a(y)$, we can apply the transformation
%function $h$ to the partition $\mathcal{P}_1$. 
If the transformation function $h$ is strictly increasing,
the application of $h$ to the partition $\mathcal{P}_1$ induces a new partition $\mathcal{P}_2 = \{x_0^2, x_1^2,\cdots, x_n^2\}$.
Otherwise, if $h$ is strictly decreasing, the application of $h$ induces a reversed partition $\mathcal{P}_2 = \{x_n^2, \cdots, x_1^2,x_0^2\}$. In the rest of this paper
we will assume that $h$ is strictly increasing, without any loss of generality.  Let 
$M_k = \sup \{f (x), h(x_{i-1}^1) = x_{i-1}^2 \leq x \leq x_i^2 = h(x_i^1)\}$, $m_k = \inf \{f (x), h(x_{i-1}^1) = x_{i-1}^2 \leq x \leq x_i^2 = h(x_i^1)\}$, and set 
\begin{equation}
\label{eq:c_up}
\mathcal{C}_U(\mathcal{P}_1,f,h) = \sum_{k=1}^n M_k \Delta x_k, 
\end{equation}
and
\begin{equation}
\label{eq:c_low}
\mathcal{C}_L(\mathcal{P}_1,f,h) = \sum_{k=1}^n m_k \Delta x_k. 
\end{equation}
The sums in Eq.~\eqref{eq:c_up} and Eq.~\eqref{eq:c_low} are respectively called the the upper and lower Cavalieri sums.
If there is a unique number $I$ that satisfies the inequality $\mathcal{C}_L(\mathcal{P}_1,f,h)\leq I \leq \mathcal{C}_U(\mathcal{P}_1,f,h)$ for all 
partitions $\mathcal{P}_1$ of $[a,b]$, then $I$ is called the Cavalieri integral of $f$ from $a(y)$ to $b(y)$ and is denoted by
\begin{equation}
\int_{a(y)}^{b(y)} f(x) dx.
\end{equation}
\end{definition}

@@

\subsection{Clarifying Example}

Consider the region $R$ bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$. This region is shown in Fig.~\ref{fig:ex1}.\\
\begin{figure}[htb]
\centering
\includegraphics[width=0.36\textwidth]{fig12}
\caption{Region bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$.}
\label{fig:ex1}
\end{figure}

\noindent
Instead of trying to determine the area of $R$ using rectangular integration strips, it seems more natural 
to use integration strips whose sides are translations of $a(y)$. Cavalieri's principle tells us that the area of an 
arbitrary shaped integration strip is the product of its height and the length of its base. The area of $R$ 
can be approximated by summing together multiple non-rectangular integration strips that inscribe $R$. This concept is illustrated in Fig.~\eqref{fig:caval2}. 
More formally, consider a partition $(x_i^1)_{i=0}^{n}$, on the $x$-axis, such that $a = x_0^1 < x_1^1 < \cdots < x_n^1 = b$, and $\Delta x_i^1 = x_{i+1}^1 - x_i^1$.
%As Fig.~\ref{fig:caval2} illustrates, the partition points $x_i^1$ are chosen to coincide with the $x$-coordinates at which the sides of the 
%integration strips intersect the $x$-axis. Fig.~\ref{fig:caval2} also shows us that another partition $(x_i^2)_{i=0}^{n}$, on the $x$-axis, such that $a' = x_0^2 < x_1^2 < \cdots < x_n^2 = b'$
%is automatically created when we make use of non-rectangular integration strips. The partition points $x_i^2$ 
%coincide with the $x$-coordinates at which the sides of the 
%integration strips intersect $f(x)$. 
Let us now construct the following sum (i.e. a lower Cavalieri sum) with which we can approximate the area 
of $R$:
\begin{equation}
\label{eq:cav_sum}
\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}
The partition points $(x_i^2)_{i=0}^{n}$ are depicted in Fig.~\ref{fig:caval2}.
The above sum can be interpreted as summing together the areas associated with $n$ non-rectangular integration strips.
The left panel of Fig.~\ref{fig:caval2} depicts the scenario in which we made use of three integration strips.  
The right panel of Fig.~\ref{fig:caval2} illustrates that the larger $n$ becomes the better Eq.~\eqref{eq:cav_sum} approximates the area of $R$. In the limit Eq.~\eqref{eq:cav_sum} approaches the Cavalieri integral
\begin{equation}
\label{eq:caval1}
\int_{a(y)}^{b(y)}f(x)\, dx = \lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}

@@

\noindent
It is however hard to evaluate Eq.~\eqref{eq:caval1} in its present form. Luckily we can transform the Cavalieri sum given in Eq.~\eqref{eq:caval1} into an ordinary Riemann sum
if we can determine an expression for $x_i^2$ in terms of the partition points $x_i^1$, for all $i=0,1,\ldots,n$. Consider the collection of functions $\{a(y) + [x_i^1-a] = x_i^2: i=0,1,\ldots,n\}$. To find the partition points $x_i^2$ in terms of $x_i^1$ we substitute the function $f(x_i^2)$ for $y$ to obtain:
\begin{align*}
a\circ f (x_i^2)+[x_i^1-1] &= x_i^2\\
x_i^2 &= \dfrac{x_i^1}{2},
\end{align*}
so that we have the general expression $x_i^2=h(x_i^1)$, with $h(x)=x/2$.\\

@@

\section{2D Geometric Interpretation}
\begin{figure}[htb]
\centering
\includegraphics[width=1\textwidth]{fig13_change}
\caption{A sum-based 2D geometric interpretation of the Riemann-Stieltjes integral. Moreover, the figure depicts the region $R$ (also see Fig.~). The integrator $g$ maps points 
in $[a',b']$ (depicted in blue) to points in $[a,b]$ (depicted in green). The function $h$ performs the reverse operation. Moreover, $g$ maps the $x$-coordinates at which
translations of $a(y)$ (depicted in red) intersect with $f(x)$ to the $x$-intercepts of the translated functions. This observation enables us to interpret the Riemann-Stieltjes integral as an infinite sum of infinitesimally small non-rectangular integration 
strips.}
\label{fig:2d_geo}
\end{figure}
%\begin{equation}
%\label{eq:cal_def_2D}
%\int_{a(y)}^{b(y)}fdx =\lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1,
%\end{equation}
%and that
%\begin{equation}
%\label{eq:rs_def_g}
%\int_{a'}^{b'} f \, dg(x) = \lim_{n\to \infty} \sum_{i=0}^{n-1} f(x_i^2)[g(x_{i+1}^2)-g(x_{i}^2)].
%\end{equation}
%Fig.~\ref{fig:2d_geo} depicts a region whose area can be determined by considering non-rectangular integration strips.
%. Note that the shape of the integration strips are determined by $a(y)$. 
%In Fig.~\ref{fig:2d_geo}, the function $a(y)$ is depicted in red. If we translate $a(y)$ along the $x$-axis we actually create two partitions on the $x$-axis, 
%i.e. $(x_i^1)_{i=0}^{n}$ and $(x_i^2)_{i=0}^{n}$ with $x_i^1\in[a,b]$ and $x_i^2\in[a',b']$.  To map partition points in $[a',b']$ to partition points 
%in $[a,b]$ we use the function $g$. The reverse is achieved through $h$. The functions $g$ and $h$ depend soly on $a(y)$ and are also 
%computed with it.
Consider the definition of the Riemann-Stieltjes integral and 
Eq.~\eqref{eq:g_int}. For the sake of continiuty we repeat Eq.~\eqref{eq:g_int} below:
\begin{equation}
\label{eq:g_int2}
\int_{a'}^{b'} f(x) dg(x) =  \lim_{n \rightarrow \infty}\sum_{i=0}^{n-1} f(x_i^2)[g(x_{i+1}^2)-g(x_{i}^2)]. 
\end{equation}
Eq.~\eqref{eq:g_int2} tells us that the function of the integrator $g$ is to map points in the interval $[a',b']$ to other 
locations on the $x$-axis. When, $g$ is monotone increasing it maps points in the interval $[a',b']$ to $[a,b]$,
with $a = g(a')$ and $b = g(b')$.\\

\noindent
To improve our understanding of $g$ lets explore a specific example of the Riemann-Stieltjes integral: 
\begin{equation}
\int_{\frac{1}{2}}^1 x~d2x. 
\end{equation}
In this example $g$ maps $[\frac{1}{2},2]$ to $[1,4]$. The mapping between these two intervals is graphically 
depicted in Fig.~\ref{fig:2d_geo}. Moreover, the interval $[\frac{1}{2},2]$ is depicted in blue,
while the interval $[1,4]$ is depicted in green.\\

\noindent
Furthermore, consider $a(y) = 1-y$ (depicted in red in Fig.~\ref{fig:2d_geo}). Interestingly, the integrator $g$ maps the $x$-coordinates where translations of $a(y)$ and $f(x)$ intersect to the $x$-intercepts of the translations of $a(y)$.
The implications of this observation is the following: 
\begin{itemize}
 \item $g(x_{i+1}^2)-g(x_{i}^2)$ represents the length of the base of a non-rectangular integration strip.
 \item $f(x_i^2)$ is the height of a non-rectangular integration strip.
 \item $f(x_i^2)[g(x_{i+1}^2)-g(x_{i}^2)]$ is the area of a non-rectangular integration strip (due to Cavalieri's principle).
 \item The Riemann-Stieltjes sum, i.e. $\sum_{i=0}^{n-1} f(x_i^2)[g(x_{i+1}^2)-g(x_{i}^2)]$, is equal to the area obtained by summing together the area of multiple 
 integration strips.
 \item In the limit the Riemann-Stieltjes sum approaches the area of $R$. 
\end{itemize}
The Riemann-Stieltjes integral, can therefore, be interpreted as the area which is obtained, by summing together the areas of an infinite number of infinitesimally small non-rectangular integration strips,
i.e. a Riemann-Stieltjes integral can be converted into an equivalent Cavalieri integral. Note, that the geometric interpretation presented
in Fig.~\ref{fig:2d_geo} is two dimensional.\\

>>>>

\title{An Alternative Geometric Interpretation of the Riemann-Stieltjes Integral}
\markright{Abbreviated Article Title}
\author{Trienko Lups Grobler}

@@

\begin{abstract}
It is desirable to have geometric interpretations for difficult mathematical ideas. In this paper we present an intuitive geometric interpretation of the Riemann-Stieltjes integral. It can be 
summarized as follows: the Riemann-Stieltjes integral can be inerpreted as an infinite sum of infinitesimally small non-rectangular 
integration strips. This interpretation is surprisingly similar to the geometric interpretation of the ordinary Riemann integral. 
It is therefore likely that students, who are encountering the Riemann-Stieltjes integral for the first time, would find this new alternative geometric interpretation useful for mastering it.
\end{abstract}

@@

\noindent
Ackermann et al. utilized the idea of using non-rectangular integration strips to perform integration
and in doing so developed the Cavalieri integral \cite{ackermann2012}. The area of a non-rectangular integration strip is computed via 
Cavalieri's principle \cite{andersen1985,eves1991, malik1984, wildberger2002, young1985}. They also showed that a Cavalieri integral can easily be converted to a Riemann-Stieltjes integral. The 
reverse operation was however not described in \cite{ackermann2012}. In this paper we address this shortcoming. This 
reverse operation actually allows us to contrive a novel geometric interpretation of the Riemann-Stieltjes integral.
This geometric interpretation is: the Riemann-Stieltjes integral can be interpreted as an infinite sum of infinitesimally small
non-rectangular integration strips. Interestingly enough, this interpretation of the Riemann-Stieltjes integral is very similar 
to the geometric interpretation of the Riemann integral. The difference being: in the case of the Riemann integral we employ 
rectangular instead of non-rectangular integration strips. This rectangular geometric interpretation has been quite useful while 
teaching the ordinary Riemann integral to undergraduate students \cite{jones2015,sealey2014}. It, therefore, seems likely that undergraduate students would find the non-rectangular geometric interpretation
of the Riemann-Stieltjes integral just as useful. It is also interesting to mention, that there are a few papers in the literature that propose to first 
teach simpler integral definitions (which are easier to grasp) before introducing students to the Riemann integral \cite{czarnocha2001,prabhu2008,ruffa2002,zaskis2014}. It therefore also seems plausible that
students will find it useful to first understand the simpler Cavalieri integral before they are exposed to the more abstract Riemann-Stieltjes integral.\\

\noindent
We first present the formal definition of the Riemann-Stieltjes and Cavalieri integrals \cite{bartle1976,ackermann2012}.
We then present the geometric interpretation proposed by G.~Bullock \cite{bullock1988}, which is followed by our alternative interpretation. We end 
the paper with some examples which aim to help the reader grasp the new alternative interpretation.

@@

\begin{definition}
Define a partition $\mathcal{P}$ of $[a',b']$ to be a set of points $x_0, x_1,\cdots,x_n$, 
where $a' = x_0 \leq x_1 \leq \cdots \leq x_n = b'$. For each partition 
$\mathcal{P}$ of $[a',b']$ write $\Delta g_k = g(x_k)-g(x_{k-1})$. Let 
$M_k = \max\{f(x),x_{k-1}\leq x \leq x_{k}\}$, $m_k = \max\{f(x),x_{k-1}\leq x \leq x_{k}\}$, and set 
\begin{equation}
\label{ref:S_up}
\mathcal{S}_U(\mathcal{P},f,g) = \sum_{k=1}^n M_k \Delta g_k, 
\end{equation}
and
\begin{equation}
\label{ref:S_low}
\mathcal{S}_L(\mathcal{P},f,g) = \sum_{k=1}^n m_k \Delta g_k. 
\end{equation}
The sums in Eq.~\eqref{ref:S_up} and Eq.~\eqref{ref:S_low} are respectively called the the upper and lower Riemann-Stieltjes sums.
If there is a unique number $I$ that satisfies the inequality $\mathcal{S}_L(\mathcal{P},f,g)\leq I \leq \mathcal{S}_U(\mathcal{P},f,g)$ for all 
partitions $P$ of $[a',b']$, then $f$ is Riemann integrable with respect to $g$ on $[a',b']$. Moreover, $I$ is called the Riemann-Stieltjes integral of $f$ from $a'$ to $b'$ and is denoted by
\begin{equation}
\int_{a'}^{b'} f(x) dg(x). 
\end{equation}
\end{definition}

@@

%\begin{theorem}
%If $a'<s<b'$, $f$ is continuous at $s$, and 
%\begin{equation}
%g(x) = 
%\begin{cases}
%0 & \textrm{if}~x\leq s\\
%1 & \textrm{if}~x > s
%\end{cases}
%\end{equation}
%then $\int_a^b f dg(x) = f(s)$.
%\end{theorem}

@@

@@

\begin{definition}[Translational Function]
A continuous real-valued function $a(y)$ is called a translational function with respect to a continuous real-valued function $f(x)$ on the interval $[a,b]$ if 
$\{x\in\mathbb{R}|a\circ f(x) + z = x\}$ is a singleton, for every $z\in[0,b-a]$ and $a(0) = a$.
\end{definition}

\begin{definition}[Transformation Function]
The mapping $h : [a, b] \rightarrow [a',b']$, which maps $x_i^1 \in [a, b]$ to $x_i^2 \in [a',b']$, is defined as
$h(x_i^1) :=$ $\{x_i^2 \in [a' ,b'] | a\circ f(x_i^2) + [x_i^1 - a] = x_i^2$ , $a = a(0)\}$
\end{definition}

@@

\begin{definition}[Inverse Transformation Function]
The mapping $g:[a′, b′] \rightarrow [a, b]$, which maps $x_i^2 \in [a' , b']$ to $x_i^1\in [a, b]$,
is defined as $g(x_i^2) := x_i^2 - a \circ f (x_i^2) + a$.
\end{definition}

@@

\begin{definition}
Define a partition $\mathcal{P}_1$ of $[a,b]$ to be a set of points $x_0, x_1,\cdots,x_n$, 
where $a = x_0^1 \leq x_1^1 \leq \cdots \leq x_n^1 = b$. For each partition 
$\mathcal{P}_1$ of $[a,b]$ write $\Delta x_k = x_k-x_{k-1}$. 
%Since both boundaries of any integration strip are necessarily
%translations of the translational function $a(y)$, we can apply the transformation
%function $h$ to the partition $\mathcal{P}_1$. 
If the transformation function $h$ is strictly increasing,
the application of $h$ to the partition $\mathcal{P}_1$ induces a new partition $\mathcal{P}_2 = \{x_0^2, x_1^2,\cdots, x_n^2\}$.
Otherwise, if $h$ is strictly decreasing, the application of $h$ induces a reversed partition $\mathcal{P}_2 = \{x_n^2, \cdots, x_1^2,x_0^2\}$. It
can be assumed that $h$ is strictly increasing, without any loss of generality.  Let 
$M_k = \sup \{f (x), h(x_{i-1}^1) = x_{i-1}^2 \leq x \leq x_i^2 = h(x_i^1)\}$, $m_k = \inf \{f (x), h(x_{i-1}^1) = x_{i-1}^2 \leq x \leq x_i^2 = h(x_i^1)\}$, and set 
\begin{equation}
\label{eq:c_up}
\mathcal{C}_U(\mathcal{P}_1,f,h) = \sum_{k=1}^n M_k \Delta x_k, 
\end{equation}
and
\begin{equation}
\label{eq:c_low}
\mathcal{C}_L(\mathcal{P}_1,f,h) = \sum_{k=1}^n m_k \Delta x_k. 
\end{equation}
The sums in Eq.~\eqref{eq:c_up} and Eq.~\eqref{eq:c_low} are respectively called the the upper and lower Cavalieri sums.
If there is a unique number $I$ that satisfies the inequality $\mathcal{C}_L(\mathcal{P}_1,f,h)\leq I \leq \mathcal{C}_U(\mathcal{P}_1,f,h)$ for all 
partitions $\mathcal{P}_1$ of $[a,b]$, then $I$ is called the Cavalieri integral of $f$ from $a(y)$ to $b(y)$ and is denoted by
\begin{equation}
\int_{a(y)}^{b(y)} f(x) dx.
\end{equation}
\end{definition}

@@

\section{Cavalieri Integral: Example}

Consider the region $R$ bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$ and $b(y)=4-y$. This region is shown in Fig.~\ref{fig:ex1}.\\
\begin{figure}[htb]
\centering
\includegraphics[width=0.36\textwidth]{fig12}
\caption{Region bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$.}
\label{fig:ex1}
\end{figure}

\noindent
Instead of trying to determine the area of $R$ using rectangular integration strips, it seems more natural 
to use integration strips whose sides are translations of $a(y)$. Cavalieri's principle tells us that the area of an 
arbitrary shaped integration strip is the product of its height and the length of its base. The area of $R$ 
can be approximated by summing together multiple non-rectangular integration strips that inscribe $R$. This concept is illustrated in Fig.~\eqref{fig:caval2}. 
More formally, consider a partition $(x_i^1)_{i=0}^{n}$, on the $x$-axis, such that $a = x_0^1 < x_1^1 < \cdots < x_n^1 = b$, and $\Delta x_i^1 = x_{i+1}^1 - x_i^1$.
%As Fig.~\ref{fig:caval2} illustrates, the partition points $x_i^1$ are chosen to coincide with the $x$-coordinates at which the sides of the 
%integration strips intersect the $x$-axis. Fig.~\ref{fig:caval2} also shows us that another partition $(x_i^2)_{i=0}^{n}$, on the $x$-axis, such that $a' = x_0^2 < x_1^2 < \cdots < x_n^2 = b'$
%is automatically created when we make use of non-rectangular integration strips. The partition points $x_i^2$ 
%coincide with the $x$-coordinates at which the sides of the 
%integration strips intersect $f(x)$. 
Let us now construct the following sum (i.e. a lower Cavalieri sum) with which we can approximate the area 
of $R$:
\begin{equation}
\label{eq:cav_sum}
\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}
The partition points $(x_i^2)_{i=0}^{n}$ are depicted in Fig.~\ref{fig:caval2}.
The above sum can be interpreted as summing together the areas associated with $n$ non-rectangular integration strips.
The left panel of Fig.~\ref{fig:caval2} depicts the scenario in which we made use of three integration strips.  
The right panel of Fig.~\ref{fig:caval2} illustrates that the larger $n$ becomes the better Eq.~\eqref{eq:cav_sum} approximates the area of $R$. In the limit Eq.~\eqref{eq:cav_sum} approaches the Cavalieri integral \cite{ackermann2012}:
\begin{equation}
\label{eq:caval1}
\int_{a(y)}^{b(y)}f(x)\, dx = \lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1.
\end{equation}

@@

\noindent
It is however hard to evaluate Eq.~\eqref{eq:caval1} in its present form. Fortuitously, we can transform the Cavalieri sum given in Eq.~\eqref{eq:caval1} into an ordinary Riemann sum
if we can determine an expression for $x_i^2$ in terms of the partition points $x_i^1$, for all $i=0,1,\ldots,n$. Consider the collection of functions $\{a(y) + [x_i^1-a] = x_i^2: i=0,1,\ldots,n\}$. To find the partition points $x_i^2$ in terms of $x_i^1$ we substitute the function $f(x_i^2)$ for $y$ to obtain:
\begin{align*}
a\circ f (x_i^2)+[x_i^1-1] &= x_i^2\\
x_i^2 &= \dfrac{x_i^1}{2},
\end{align*}
so that we have the general expression $x_i^2=h(x_i^1)$, with $h(x)=x/2$.\\

@@

\section{Existing Geometric Interpretation}
G.~Bullock proposed one possible geometric interpretation of the Riemann-Stieltjes integral \cite{bullock1988}. 
The basic premise behind his approach is to represent $f(x)$ and $g(x)$ as three-dimensional sheets. 
The three dimensional coordinate system which G.~Bullock proposes to realize this representation is depicted in 
Fig.~\ref{fig:bullock}. The $f$-sheet is constructed by simply extending the curve $f$ which resides in the $fx$-plane 
along the $g$-direction. The $g$-sheet is constructed similarly. Note that, since $f$ is independent of $g$, the $f$-sheet remains 
constant along the $g$-direction. Similarly, the $g$-sheet remains constant along the $f$-direction. 
A new sheet, known simply as the ``fence'', can be created by chopping off any part of the $g$-sheet that protrudes above the 
$f$-sheet and below the $gx$-plane. This sheet can now be projected onto the $fg$-plane. The area of the region that is formed 
via this projection can be interpreted as the value which is obtained when the Riemann-Stieltjes integral is evaluated.\\

\noindent
Let us now try to make the above ideas more concrete by applying them to an example. Let $f(x)=x^2$ and 
$g(x)=x$. The sheets associated with these two functions are depicted in Fig.~\ref{fig:bullock}. 
The $f$-sheet is depicted in blue, while the $g$-sheet is red. The ``fence'' is depicted in green. The region
obtained by projecting the ``fence'' on to the $fg$-plane is depicted in black.\\

\noindent
\begin{remark}
It is interesting to note that G.~Bullock follows a similar route than we do in terms of developing his geometrical framework.
He starts of with quite strict requirements for $f$ ang $g$ and relaxes more and more of these requirements in his latter 
examples. 
\end{remark}

\begin{figure}[htb]
\centering
\includegraphics[width=1.0\textwidth]{3D_final-crop.pdf}
\caption{An example of using G.~Bullocks' geometric framework to interpret the Riemann-Stieltjes integral. In this example 
$f(x)$ is represented by the blue sheet, $g(x)$ is represented by the red sheet. Moreover, $f(x)=x^2$ and $g(x)=x$. The green sheet, simply known as the ``fence'', is formed 
by chopping of any part of the $g$-sheet which protrudes above the $f$-sheet and below the $gx$-plane. Projecting the ``fence''
onto the $fg$-plane produces the visible gray region in the $fg$-plane. The area of this gray region in the $fg$-plane is equal to the Riemann-Stieltjes integral:
$\int_{0}^{3} f(x) dg(x)$.}
\label{fig:bullock}
\end{figure}

\section{Alternative Geometric Interpretation}
\begin{figure}[htb]
\centering
\includegraphics[width=1\textwidth]{fig13_change}
\caption{A non-rectangular sum-based geometric interpretation of the Riemann-Stieltjes integral. Moreover, the figure depicts the region $R$ (also see Fig.~). The integrator $g$ maps points 
in $[a',b']$ (depicted in blue) to points in $[a,b]$ (depicted in green). The function $h$ performs the reverse operation. Moreover, $g$ maps the $x$-coordinates at which
translations of $a(y)$ (depicted in red) intersect with $f(x)$ to the $x$-intercepts of the translated functions. This observation enables us to interpret the Riemann-Stieltjes integral as an infinite sum of infinitesimally small non-rectangular integration 
strips.}
\label{fig:2d_geo}
\end{figure}
%\begin{equation}
%\label{eq:cal_def_2D}
%\int_{a(y)}^{b(y)}fdx =\lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1,
%\end{equation}
%and that
%\begin{equation}
%\label{eq:rs_def_g}
%\int_{a'}^{b'} f \, dg(x) = \lim_{n\to \infty} \sum_{i=0}^{n-1} f(x_i^2)[g(x_{i+1}^2)-g(x_{i}^2)].
%\end{equation}
%Fig.~\ref{fig:2d_geo} depicts a region whose area can be determined by considering non-rectangular integration strips.
%. Note that the shape of the integration strips are determined by $a(y)$. 
%In Fig.~\ref{fig:2d_geo}, the function $a(y)$ is depicted in red. If we translate $a(y)$ along the $x$-axis we actually create two partitions on the $x$-axis, 
%i.e. $(x_i^1)_{i=0}^{n}$ and $(x_i^2)_{i=0}^{n}$ with $x_i^1\in[a,b]$ and $x_i^2\in[a',b']$.  To map partition points in $[a',b']$ to partition points 
%in $[a,b]$ we use the function $g$. The reverse is achieved through $h$. The functions $g$ and $h$ depend soly on $a(y)$ and are also 
%computed with it.
Consider the definition of the Riemann-Stieltjes integral and 
Eq.~\eqref{eq:g_int}. For the sake of continiuty we repeat Eq.~\eqref{eq:g_int} below:
\begin{equation}
\label{eq:g_int2}
\int_{a'}^{b'} f(x) dg(x) =  \lim_{n \rightarrow \infty}\sum_{i=0}^{n-1} f(x_i^2)[g(x_{i+1}^2)-g(x_{i}^2)], 
\end{equation}
with $(x_i^2)_{i=0}^n$ being arbitrary partition points on the $x$-axis.
Eq.~\eqref{eq:g_int2} tells us that the function of the integrator $g$ is to map points in the interval $[a',b']$ to other 
locations on the $x$-axis. When, $g$ is monotone increasing it maps points in the interval $[a',b']$ to $[a,b]$,
with $a = g(a')$ and $b = g(b')$.\\
